Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates

### 2025-05-23 performance_comparison.tex
[MM] few changes

Figure~\ref{fig:ol_telemetries} presents the variation of four signals with respect to the not-robust trajectory, for the three previously mentioned cases. The panels display the longitudinal speed $u$ (first panel), wheel steering angle $\de$ (second panel), total longitudinal force $F_x$ (third panel), and lateral deviation of the center of mass from the track centerline $e$ (fourth panel). Except for the $u$ panel, each plot includes dashed black lines that indicate the sign of the reference signal-whether it is positive, negative, or numerically zero. These sign indicators are represented as three distinct levels---high, zero, and low---corresponding to positive, zero, and negative values of the reference signal, respectively. The levels are scaled appropriately in each plot to ensure clarity of visualization.

\begin{figure}
	\centering
	\includegraphics{Fig/ol_telemetries.pdf}
	\caption{Variation of four signals w.r.t. the not-robust trajectory for the three configurations TLC (green line), FLC (red line) and TLC+FLC (blue line). The panels display the longitudinal speed $u$ (first panel), wheel steering angle $\de$ (second panel), total longitudinal force $F_x$ (third panel), and lateral deviation of the center of mass from the track centerline $e$ (fourth panel), over the curvilinear abscissa $\al$. The dashed lines in the last three panels indicate the sign of the reference signal, allowing one to determine whether a positive variation with respect to the baseline corresponds to an increase or a decrease in the absolute value. This distinction is particularly important for quantities that depend on the direction of the turn, such as the wheel steering angle $\de$ and the lateral deviation $e$, while in the case of $F_x$, it indicates whether the vehicle is accelerating or braking.}
	\label{fig:ol_telemetries}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics{Fig/ol_saturation.pdf}
	\caption{Tyre usage for front axle (first panel) and rear axle (second panel) over curvilinear abscissa $\al$. The plotted lines represent the quantity $\left[ \left( \frac{X_j(\bx,\bu)}{\mu_{x,j}} \right)^2+ \left( \frac{Y_j(\bx,\bu)}{\mu_{y,j}}\right)^2\right]\frac{1}{Z_j^2(\bx,\bu)}$, where $j=1$ denotes the front axle and $j=2$ the rear axle. The dashed black line corresponds to the reference (not-robust) configuration, while the green, red, and blue lines represent the robust configurations TLC, FLC, and TLC+FLC, respectively.}
	\label{fig:ol_saturation}
\end{figure}

\subsection{Simulations with noise realization}
This analysis aims to provide empirical validation of the robust trajectories obtained using the methods proposed in this work. For conciseness, the comparison is limited to a nominal trajectory and a trajectory computed using the closed-loop method with only the track limit constraint robustified.
An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented along the not-robust trajectory. In contrast, the robust trajectory is tracked using the controller directly obtained from the optimization process.

>>>>

Figure~\ref{fig:ol_telemetries} presents the variation of four signals with respect to the nominal trajectory, for the three previously mentioned cases. The panels display the longitudinal speed $u$ (first panel), wheel steering angle $\de$ (second panel), total longitudinal force $F_x$ (third panel), and lateral deviation of the center of mass from the track centerline $e$ (fourth panel). Except for the $u$ panel, each plot includes dashed black lines that indicate the sign of the reference signal-whether it is positive, negative, or numerically zero. These sign indicators are represented as three distinct levels---high, zero, and low---corresponding to positive, zero, and negative values of the reference signal, respectively. The levels are scaled appropriately in each plot to ensure clarity of visualization.

\begin{figure}
	\centering
	\includegraphics{Fig/ol_telemetries.pdf}
	\caption{Evolution of the longitudinal speed u, wheel steering angle $\de$, longitudinal force $F_x$, and lateral deviation $e$ are reported for three robustification scenarios: TLC (Track Limit Constraints), FLC (Friction Limit Constraint), and TLC+FLC (both constraints enforced concurrently). Each profile is shown relative to the baseline (nominal) trajectory to highlight the effect of each constraint-handling strategy. The dashed lines in the last three panels indicate the sign of the reference signal, allowing one to determine whether a positive variation with respect to the baseline corresponds to an increase or a decrease in the absolute value. This distinction is particularly important for quantities that depend on the direction of the turn, such as the wheel steering angle $\de$ and the lateral deviation $e$, while in the case of $F_x$, it indicates whether the vehicle is accelerating or braking.}
	\label{fig:ol_telemetries}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics{Fig/ol_saturation.pdf}
	\caption{Tire usage for front axle (first panel) and rear axle (second panel) over curvilinear abscissa $\al$. The plotted lines represent the quantity $\left[ \left( \frac{X_j(\bx,\bu)}{\mu_{x,j}} \right)^2+ \left( \frac{Y_j(\bx,\bu)}{\mu_{y,j}}\right)^2\right]\frac{1}{Z_j^2(\bx,\bu)}$, where $j=1$ denotes the front axle and $j=2$ the rear axle. The dashed black line corresponds to the reference (nominal) configuration, while the green, red, and blue lines represent the robust configurations TLC, FLC, and TLC+FLC, respectively.}
	\label{fig:ol_saturation}
\end{figure}

\subsection{Simulations with noise realization}
This analysis aims to provide empirical validation of the robust trajectories obtained using the methods proposed in this work. For conciseness, the comparison is limited to a nominal trajectory and a trajectory computed using the closed-loop method with only the track limit constraint robustified.
An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented along the nominal trajectory. In contrast, the robust trajectory is tracked using the controller directly obtained from the optimization process.

### 2025-05-23 stochastic_vehicle_dynamics.tex
[MM] some constraint references were not updated
[MM] few changes

\subsection{Robust friction limit constraint formulation}
\label{sec:FLC}
The model employs a tire formulation that neglects combined slip effects, i.e., it does not account for the simultaneous utilization of longitudinal and lateral tyre forces. As a result, the optimal control problem must include an additional constraint to ensure that the total ground reaction forces remain within the bounds of the tyre's adherence ellipse.

@@

\subsection{Robust track limit constraint formulation}
\label{sec:TLC}
An on-track constraint is required to ensure that the vehicle remains within the circuit boundaries. The formulation adopted introduces an algebraic variable $e$, which satisfies the following constraint:
\begin{equation}
	\begin{Bmatrix}
		x\\y
	\end{Bmatrix}
	-\boldsymbol{c} - e\bn=\boldsymbol{0} \label{eq:onplane_constraint}
\end{equation}
%\begin{equation}
%	\bp-\boldsymbol{c}_k - e_k\bn_k=\boldsymbol{0} \label{eq:onplane_constraint}
%\end{equation}
where $\boldsymbol{c}$ indicates the position of the centerline, and $\bn$ is the unit normal vector to the centerline. Equation~\eqref{eq:onplane_constraint} constraints the CoM of the vehicle to lie on the vertical plane defined by the unit normal vector $\bn$. The previously introduced algebraic variable $e$ is further constrained to take values between $e_\textrm{min}$ and $e_\textrm{max}$, which are determined based on the width of the circuit and the vehicle's front and rear tracks.

>>>>

\subsection{Robust friction limit constraint formulation}
\label{sec:FLC}
The model employs a tire formulation that neglects combined slip effects, i.e., it does not account for the simultaneous utilization of longitudinal and lateral tire forces. As a result, the optimal control problem must include an additional constraint to ensure that the total ground reaction forces remain within the bounds of the tire's adherence ellipse.

@@

\subsection{Robust track limit constraint formulation}
\label{sec:TLC}
An track limit constraint is required to ensure that the vehicle remains within the circuit boundaries. The formulation adopted introduces an algebraic variable $e$, which satisfies the following constraint:
\begin{equation}
	\begin{Bmatrix}
		x\\y
	\end{Bmatrix}
	-\boldsymbol{c} - e\bn=\boldsymbol{0} \label{eq:onplane_constraint}
\end{equation}
%\begin{equation}
%	\bp-\boldsymbol{c}_k - e_k\bn_k=\boldsymbol{0} \label{eq:onplane_constraint}
%\end{equation}
where $\boldsymbol{c}$ indicates the position of the centerline, and $\bn$ is the unit normal vector to the centerline. Equation~\eqref{eq:onplane_constraint} constraints the CoM of the vehicle to lie on the vertical plane defined by the unit normal vector $\bn$. The previously introduced algebraic variable $e$ is further constrained to take values between $e_\textrm{min}$ and $e_\textrm{max}$, which are determined based on the width of the circuit and the vehicle's front and rear tracks.
